---
title: "Query OSM"
author: "Brian Yandell"
date: "2024-09-17"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Use R to query OpenStreetMap data for colleges and universities in the US
using the Overpass API.
Query times out if we try the whole country, but works for a state.
Campuses are either in `osm_polygons` or `osm_multipolygons` and have full names
in column `name` and possible shorter names in `alt_name`, `alt_name_1` and
`alt_name_2`.
Probably rely on `alt_name` and combine polygons and multipolygons list
for selected state.

Next thing to do is try to find TCUs.
Need to use another list to locate them.

```{r eval=FALSE}
install.packages("httr")
install.packages("jsonlite")
library(httr)
library(jsonlite)
```

Define the Overpass API query:

```{r}
query <- osmdata::opq(bbox = "Wisconsin") |>
  osmdata::add_osm_feature(key = "amenity", value = c("university", "college"))
```

```{r}
osm_data <- osmdata::osmdata_sf(query)
```

```{r}
polys <- osm_data$osm_polygons |>
  dplyr::select(-geometry) |>
  tibble::as_tibble() |>
  dplyr::select(alt_name, alt_name_1, alt_name_2, name) |>
  tidyr::pivot_longer(dplyr::everything(),
                      names_to = "source", values_to = "name") |>
  dplyr::filter(!is.na(name), name != "")
```

```{r}
multis <- osm_data$osm_multipolygons |>
  dplyr::select(-geometry) |>
  tibble::as_tibble() |>
  dplyr::select(alt_name, alt_name_1, alt_name_2, name) |>
  tidyr::pivot_longer(dplyr::everything(),
                      names_to = "source", values_to = "name") |>
  dplyr::filter(!is.na(name), name != "")
```

```{r}
m <- match(polys$name, multis$name)
mm <- match(multis$name, polys$name)
polys$name[!is.na(m)]
multis$name[!is.na(mm)]
```
```{r}
sum(multis$source == "alt_name")
sum(polys$source == "alt_name")
```

```{r}
polys |> dplyr::filter(grepl("Menominee", name))
```

```{r}
dplyr::filter(osm_data$osm_polygons, grepl("Menominee", name)) |>
  dplyr::select(name, operator)
```


```{r}
query <- '
[out:json][timeout:25];
area["name"="California"][admin_level=2];
(
  node"amenity"="university";
  way"amenity"="university";
  relation"amenity"="university";
  node"amenity"="college";
  way"amenity"="college";
  relation"amenity"="college";
);
out body;
>;
out skel qt;
'
```

Send the query to the Overpass API:

```{r}
response <- httr::POST(
  url = "http://overpass-api.de/api/interpreter",
  body = list(data = query),
  encode = "form"
)
```

Parse the response:

```{r}
data <- jsonlite::fromJSON(httr::content(response, "text", encoding = "UTF-8"))
```

# Extract nodes, ways, and relations

```{r}
nodes <- data$elements[data$elements$type == "node", ]
ways <- data$elements[data$elements$type == "way", ]
relations <- data$elements[data$elements$type == "relation", ]
```

View the first few entries

```{r}
head(nodes)
head(ways)
head(relations)
```
